---
title: 'Subset Test: non-APOE4'
author: "Duruo Li (Dora)"
date: "9/23/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```
```{r, include=FALSE}
library("knitr")
library(mice)
library(lattice)
library(cobalt)
library(sandwich)
library(lmtest)
library(dplyr)
library(ggplot2)
library(texreg)
library(car)
```

## 0. Input
```{r}
df0<-read.csv("./data/data1_1.csv")
```

```{r}
head(df0)
cate_names<-c("SexCode","Race","ApoE_Status", "APOE",  "apoE4dose", "Cognitive_Impairment_Binary_Score")
df0[cate_names]<-lapply(df0[cate_names], factor)
#str(df0)
```

```{r}
names(df0)[names(df0) == "ï..ID"] <- 'ID'
names<-colnames(df0)
#names
control_names<-names[3:13]
idp_names<-names[16:38]
#control_names
idp_names
```
```{r}
df_noE4 <- df0 %>%
  filter(ApoE_Status==0)
df_E4 <- df0 %>%
  filter(ApoE_Status==1)
```

Missing pattern:
```{r}
df1<-df_noE4[c("ID", 'Amyloid_Status',control_names,idp_names)]
missing_names<-colnames(df1)[apply(is.na(df1), 2, any)]

md.pattern(df1[,missing_names], rotate.names = TRUE)
colSums(is.na(df1[,missing_names])) #number of missing values
#rlv_names
```
## 1. E4[-]: compare AD[+] vs AD[-]
### 1.1 Visualization
```{r}
# Create a function to plot histograms
plot_histograms <- function(var_name) {
  p <- ggplot(df, aes_string(x = var_name, fill = "factor(Amyloid_Status)")) +
    geom_histogram(alpha = 0.5, position = "identity", bins = 30, aes(y = ..density..), color = "white") +
    labs(x = var_name, fill = "Amyloid Status") +
    scale_fill_manual(values = c("red", "blue")) +
    theme_minimal()
  
  print(p)
}

# Create a function to plot KDEs
plot_kdes <- function(var_name) {
  p <- ggplot(df, aes_string(x = var_name, color = "factor(Amyloid_Status)")) +
    geom_density(size = 1.5) +
    labs(x = var_name, color = "Amyloid Status") +
    theme_minimal()
  
  print(p)
}

# Loop through each variable in idp_names to create and display plots
for (var_name in idp_names) {
  df <- df_noE4
  plot_histograms(var_name)
  plot_kdes(var_name)
}
```


### 1.2 regression
In this subset, if only consider independent variables and Amyloid_Status, there is only one row containing missing values, so no need to use imputed data.
```{r}
# Assume df0 and idp_names are already defined
df_noE4 <- df0 %>%
  filter(ApoE_Status == 0)

# Initialize an empty list to store significant models
significant_models <- list()

# Loop through each variable in idp_names
for (idp in idp_names) {
  
  # Build the formula for lm
  formula <- as.formula(paste(idp, "~ Amyloid_Status"))
  
  # Fit the linear model
  model <- lm(formula, data = df_noE4)
  
  # Check the significance of the Amyloid_Status coefficient
  coef_summary <- summary(model)$coefficients
  
  # The row corresponding to Amyloid_Status in coef_summary will be 2
  # Check if the p-value is less than 0.1
  if (coef_summary[2, 4] <= 0.1) {
    significant_models[[idp]] <- model
  }
}

# Now significant_models contains all models with a significant coefficient 
# for Amyloid_Status at a level of 0.1

```
```{r}
significant_models
summary(significant_models[[1]])#p=0.134
```
No p<0.1 (no difference between [+] and [-] group)

### 1.3 imbalance test
```{r}
formula <- as.formula(paste("Amyloid_Status", "~", paste(idp_names, collapse = " + ")))
balance_table<-bal.tab(formula, data = df_noE4, disp = c("mean", "sd"), thresholds = 0.1)
#(35.7712-35.2573)/sqrt((7.7709^2+9.6992^2)/2)
print(balance_table)
```
```{r}

# Identify unbalanced variables
unbalanced <- abs(balance_table$Balance$Diff.Un) > 0.1

# Extract names of unbalanced variables
unbalanced_var_names <- rownames(balance_table$Balance)[unbalanced]

# Print or return unbalanced variable names
print(unbalanced_var_names)
```
## 1. E4[+]: Idp ~ Amyloid_Status
### 2.1 Visualization
```{r}
# Create a function to plot histograms
plot_histograms <- function(var_name) {
  p <- ggplot(df, aes_string(x = var_name, fill = "factor(Amyloid_Status)")) +
    geom_histogram(alpha = 0.5, position = "identity", bins = 30, aes(y = ..density..), color = "white") +
    labs(x = var_name, fill = "Amyloid Status") +
    scale_fill_manual(values = c("red", "blue")) +
    theme_minimal()
  
  print(p)
}

# Create a function to plot KDEs
plot_kdes <- function(var_name) {
  p <- ggplot(df, aes_string(x = var_name, color = "factor(Amyloid_Status)")) +
    geom_density(size = 1.5) +
    labs(x = var_name, color = "Amyloid Status") +
    theme_minimal()
  
  print(p)
}

# Loop through each variable in idp_names to create and display plots
for (var_name in idp_names) {
  df <- df_E4
  plot_histograms(var_name)
  plot_kdes(var_name)
}
```


### 2.2 regression
In this subset, if only consider independent variables and Amyloid_Status, there is only one row containing missing values, so no need to use imputed data.
```{r}
# Initialize an empty list to store significant models
significant_models <- list()

# Loop through each variable in idp_names
for (idp in idp_names) {
  
  # Build the formula for lm
  formula <- as.formula(paste(idp, "~ Amyloid_Status"))
  
  # Fit the linear model
  model <- lm(formula, data = df_E4)
  
  # Check the significance of the Amyloid_Status coefficient
  coef_summary <- summary(model)$coefficients
  
  # The row corresponding to Amyloid_Status in coef_summary will be 2
  # Check if the p-value is less than 0.1
  if (coef_summary[2, 4] <= 0.1) {
    significant_models[[idp]] <- model
  }
}

# Now significant_models contains all models with a significant coefficient 
# for Amyloid_Status at a level of 0.1

```
```{r}
significant_models
#summary(significant_models[[1]])#p=0.134
```
No p<0.1 (no difference between [+] and [-] group)

### 2.3 imbalance test
```{r}
formula <- as.formula(paste("Amyloid_Status", "~", paste(idp_names, collapse = " + ")))
balance_table<-bal.tab(formula, data = df_E4, disp = c("mean", "sd"), thresholds = 0.1)
#(35.7712-35.2573)/sqrt((7.7709^2+9.6992^2)/2)
print(balance_table)
```
```{r}
# Identify unbalanced variables
unbalanced <- abs(balance_table$Balance$Diff.Un) > 0.1

# Extract names of unbalanced variables
unbalanced_var_names <- rownames(balance_table$Balance)[unbalanced]

# Print or return unbalanced variable names
print(unbalanced_var_names)
```

## 3. Visualization: Entire vs E4+ vs E4-
```{r}
# Modify the functions to accept a title argument
plot_histograms <- function(var_name, df, title) {
  p <- ggplot(df, aes_string(x = var_name, fill = "factor(Amyloid_Status)")) +
    geom_histogram(alpha = 0.5, position = "identity", bins = 30, aes(y = ..density..), color = "white") +
    labs(x = var_name, fill = "Amyloid Status") +
    scale_fill_manual(values = c("red", "blue")) +
    theme_minimal() +
    ggtitle(title)  # Add title to the plot
  
  print(p)
}

plot_kdes <- function(var_name, df, title) {
  p <- ggplot(df, aes_string(x = var_name, color = "factor(Amyloid_Status)")) +
    geom_density(size = 1.5) +
    labs(x = var_name, color = "Amyloid Status") +
    theme_minimal() +
    ggtitle(title)  # Add title to the plot
  
  print(p)
}

# Specify titles for each dataframe
title_df_noE4 <- "Non-APOE4: AD+ vs AD-"
title_df_E4 <- "APOE4: AD+ vs AD-"
title_df1 <- "Entire Data: AD+ vs AD-"

# Loop through each variable in idp_names to create and display plots
for (var_name in idp_names) {
  plot_histograms(var_name, df=df_noE4, title=title_df_noE4)
  plot_histograms(var_name, df=df_E4, title=title_df_E4)
  plot_histograms(var_name, df=df0, title=title_df1)
  plot_kdes(var_name, df=df_noE4, title=title_df_noE4)
  plot_kdes(var_name, df=df_E4, title=title_df_E4)
  plot_kdes(var_name, df=df0, title=title_df1)
}

```
## Appendix: ***Tests based on Domain Knowledge
Get imputed data:
```{r}
num_imputations <- 5
imp_pmm <- mice(data=df1, m = num_imputations, method="pmm", printFlag = FALSE, seed = 1)
imp_dfs_long <- complete(imp_pmm, action = "long", include = TRUE)
imp_dfs_sep<-split(imp_dfs_long, imp_dfs_long$.imp)
```
```{r}
imp_dfs_sep
```

Regression: Amyloid Status ~ independent combinations + control combinations
```{r}
# Assume imp_dfs_sep is already defined
all_models <- list()
significant_models <- list()

# Define the combinations of control and independent variables
control_combs <- list(
  c("Hippocampal_Vol", "SexCode"),
  c("Hippocampal_Vol",  "SexCode", "Cognitive_Impairment_Binary_Score"),
  c("Hippocampal_Vol",  "SexCode", "Cognitive_Impairment_Binary_Score", "BMI"),
  c("Hippocampal_Vol",  "SexCode", "Cognitive_Impairment_Binary_Score", "BMI", "Age")
)

idp_combs <- list(
  c("cct_sAPPb", "pdr_sAPPb"),
  c("cct_sAPPa", "cct_sAPPb", "cct_ratio_sAPPb_sAPPa", "pdr_sAPPa", "pdr_sAPPb", "pdr_ratio_sAPPb_sAPPa"),
  c("cct_sAPPb", "cct_ratio_sAPPb_totalAb", "pdr_sAPPb", "pdr_ratio_sAPPb_totalAb"),
  c("FTR_ratio_sAPPb_sAPPa", "FCR_ratio_sAPPb_sAPPa", "dlt_ratio_sAPPb_sAPPa"),
  c("FTR_sAPPb", "FTR_ratio_sAPPb_Ab40", "FCR_sAPPb", "FCR_ratio_sAPPb_Ab40", "dlt_sAPPb", "dlt_ratio_sAPPb_Ab40"),
  c("FTR_sAPPa", "FTR_sAPPb", "FTR_ratio_sAPPb_sAPPa", "FCR_sAPPa", "FCR_sAPPb", "FCR_ratio_sAPPb_sAPPa", "dlt_sAPPa", "dlt_sAPPb", "dlt_ratio_sAPPb_sAPPa")
)

# Iterate through all combinations
for(i in seq_along(control_combs)) {
  for(j in seq_along(idp_combs)) {
    
    # Create the formula string
    formula_str <- paste(
      "Amyloid_Status ~", 
      paste(control_combs[[i]], collapse = " + "), 
      "+", 
      paste(idp_combs[[j]], collapse = " + ")
    )
    
    # Convert to formula object
    formula <- as.formula(formula_str)
    
    # Fit models to each imputed dataset
    imputed_models <- lapply(imp_dfs_sep, function(df) {
      df_noE4 <- df %>% filter(ApoE_Status==0)
      lm(formula = formula, family = binomial, data = df_noE4)
    })
    
    # Pool the models
    pooled_model <- pool(imputed_models)
    
    # Store all models
    all_models[[paste(i, j, sep = "_")]] <- pooled_model
    
    # Check for significance of any independent variable
    coef_summary <- summary(pooled_model)$coefficients
    if(any(coef_summary[-(1:length(control_combs[[i]])), 4] < 0.1)) {
      significant_models[[paste(i, j, sep = "_")]] <- pooled_model
    }
  }
}

# Now all_models contains all 30 models, and significant_models contains those 
# with at least one significant independent variable at a level of 0.1.

```
```{r}
significant_models
```
