---
title: 'Subset Test: non-APOE4'
author: "Duruo Li (Dora)"
date: "9/23/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```
```{r, include=FALSE}
library(mice)
library(lattice)
library(cobalt)
library(sandwich)
library(lmtest)
library(dplyr)
library(ggplot2)
library(texreg)
library(car)
```

## 0. Input
```{r}
df0<-read.csv("./data/data1_1.csv")
```

```{r}
head(df0)
cate_names<-c("SexCode","Race","ApoE_Status", "APOE",  "apoE4dose", "Cognitive_Impairment_Binary_Score")
df0[cate_names]<-lapply(df0[cate_names], factor)
#str(df0)
```

```{r}
names(df0)[names(df0) == "ï..ID"] <- 'ID'
names<-colnames(df0)
#names
control_names<-names[3:13]
idp_names<-names[16:38]
#control_names
idp_names
```
```{r}
df_AD <- df0 %>%
  filter(Amyloid_Status==1)
df_noAD <- df0 %>%
  filter(Amyloid_Status==0)
```

Missing pattern:
```{r}
df1<-df_noE4[c("ID", 'Amyloid_Status',control_names,idp_names)]
missing_names<-colnames(df1)[apply(is.na(df1), 2, any)]

md.pattern(df1[,missing_names], rotate.names = TRUE)
colSums(is.na(df1[,missing_names])) #number of missing values
#rlv_names
```

## 1. AD[+]: compare nonAPOE4 vs APOE4
### 1.1 Visualization
```{r}
# Modify the functions to accept a title argument
plot_histograms <- function(var_name, df, title) {
  p <- ggplot(df, aes_string(x = var_name, fill = "factor(ApoE_Status)")) +
    geom_histogram(alpha = 0.5, position = "identity", bins = 30, aes(y = ..density..), color = "white") +
    labs(x = var_name, fill = "APOE Status") +
    scale_fill_manual(values = c("red", "blue")) +
    theme_minimal() +
    ggtitle(title)  # Add title to the plot
  
  print(p)
}

plot_kdes <- function(var_name, df, title) {
  p <- ggplot(df, aes_string(x = var_name, color = "factor(ApoE_Status)")) +
    geom_density(size = 1.5) +
    labs(x = var_name, color = "ApoE Status") +
    theme_minimal() +
    ggtitle(title)  # Add title to the plot
  
  print(p)
}

df_AD <- df0 %>%
  filter(Amyloid_Status==1)
for (var_name in idp_names) {
  plot_histograms(var_name, df=df_AD, title="AD[+]: E4+ vs E4-")
  plot_kdes(var_name, df=df_AD, title="AD[+]: E4+ vs E4-")
}
```
```{r}
df_AD %>%
  group_by(ApoE_Status) %>%
  summarise(n=n())
```
```{r}
formula <- as.formula(paste("ApoE_Status", "~", paste(idp_names, collapse = " + ")))
balance_table<-bal.tab(formula, data = df_AD, disp = c("mean", "sd"), thresholds = 0.1)
print(balance_table)
```
Missing pattern:
```{r}
df1<-df_AD[c("ID", 'Amyloid_Status',control_names,idp_names)]
missing_names<-colnames(df1)[apply(is.na(df1), 2, any)]

md.pattern(df1[,missing_names], rotate.names = TRUE)
colSums(is.na(df1[,missing_names])) #number of missing values
#rlv_names
```

### 1.2 regression: Idp ~ ApoE_Status
```{r}
# Initialize an empty list to store significant models
significant_models <- list()

# Loop through each variable in idp_names
for (idp in idp_names) {
  
  # Build the formula for lm
  formula <- as.formula(paste(idp, "~ ApoE_Status"))
  
  # Fit the linear model
  model <- lm(formula, data = df_AD)
  
  # Check the significance of the Amyloid_Status coefficient
  coef_summary <- summary(model)$coefficients
  
  # The row corresponding to Amyloid_Status in coef_summary will be 2
  # Check if the p-value is less than 0.1
  if (coef_summary[2, 4] <= 0.1) {
    significant_models[[idp]] <- model
  }
}

# Now significant_models contains all models with a significant coefficient 
# for Amyloid_Status at a level of 0.1

```
```{r}
significant_models
summary(significant_models[[1]])#p=0.134
```
```{r}
model <- lm(dlt_ratio_sAPPb_sAPPa~ApoE_Status, data = df_AD)
summary(model)
```
### Oversampling?
```{r}
df_AD %>%
  group_by(ApoE_Status) %>%
  summarise(n=n())
```
```{r}
df_AD

```

### Multiple comparisons: bonferroni correction

```{r}
# Initialize an empty list to store p-values
p_values <- vector()

# Loop through each variable in idp_names
for (idp in idp_names) {
  
  # Build the formula for lm
  formula <- as.formula(paste(idp, "~ ApoE_Status"))
  
  # Fit the linear model
  model <- lm(formula, data = df_AD)
  
  # Check the significance of the ApoE_Status coefficient
  coef_summary <- summary(model)$coefficients
  
  # The row corresponding to ApoE_Status in coef_summary will be 2
  # Store the p-value in p_values
  p_values <- c(p_values, coef_summary[2, 4])
}

# Now adjust the p-values using Holm-Bonferroni method
adjusted_p_values <- p.adjust(p_values, method = "BH")

# Initialize an empty list to store significant models
significant_models <- list()

# Loop through the adjusted p-values and check if they are below 0.1
for (i in seq_along(adjusted_p_values)) {
  if (adjusted_p_values[i] <= 0.1) {
    
    # Build the formula for lm
    formula <- as.formula(paste(idp_names[i], "~ ApoE_Status"))
    
    # Fit the linear model
    model <- lm(formula, data = df_AD)
    
    # Store the model in significant_models
    significant_models[[idp_names[i]]] <- model
  }
}

# Now significant_models contains all models with a significant coefficient 
# for ApoE_Status at a level of 0.1, adjusted for multiple comparisons

```
```{r}
adjusted_p_values
```

```{r}
significant_models
summary(significant_models[[1]])#p=0.134
```
### Imbalance
```{r}
formula <- as.formula(paste("ApoE_Status", "~", paste(control_names, collapse = " + ")))
balance_table<-bal.tab(formula, data = df_AD, disp = c("mean", "sd"), thresholds = 0.1)
#(35.7712-35.2573)/sqrt((7.7709^2+9.6992^2)/2)
print(balance_table)
```

```{r}
# Initialize an empty list to store p-values
p_values <- vector()

# Loop through each variable in idp_names
for (idp in idp_names) {
  
  # Build the formula for lm
  formula <- as.formula(paste(idp, "~ ApoE_Status + SexCode + BMI + CDR"))
  
  # Fit the linear model
  model <- lm(formula, data = df_AD)
  
  # Check the significance of the ApoE_Status coefficient
  coef_summary <- summary(model)$coefficients
  
  # The row corresponding to ApoE_Status in coef_summary will be 2
  # Store the p-value in p_values
  p_values <- c(p_values, coef_summary[2, 4])
}

# Now adjust the p-values using Holm-Bonferroni method
adjusted_p_values <- p.adjust(p_values, method = "holm")

# Initialize an empty list to store significant models
significant_models <- list()

# Loop through the adjusted p-values and check if they are below 0.1
for (i in seq_along(adjusted_p_values)) {
  if (adjusted_p_values[i] <= 0.1) {
    
    # Build the formula for lm
    formula <- as.formula(paste(idp_names[i], "~ ApoE_Status + SexCode + BMI + CDR"))
    
    # Fit the linear model
    model <- lm(formula, data = df_AD)
    
    # Store the model in significant_models
    significant_models[[idp_names[i]]] <- model
  }
}

# Now significant_models contains all models with a significant coefficient 
# for ApoE_Status at a level of 0.1, adjusted for multiple comparisons

```
```{r}
adjusted_p_values
```
### Association between control vars and outcomes
```{r}
control_names
```
```{r}
cont_control_names<-c("Age", "BMI", "CDR", "CDR_SOB", "Hippocampal_Vol")
bi_control_names<-c("SexCode", "Race", "Cognitive_Impairment_Binary_Score")
```

```{r}
# Load necessary library
library(reshape2)
library(ggplot2)

# Assume df is your data frame
df <- df_AD  # replace 'your_dataframe' with the actual name of your data frame

# Initialize empty matrices to store correlation coefficients
var_names<-cont_control_names
pearson_matrix <- matrix(NA, nrow = length(idp_names), ncol = length(var_names),
                         dimnames = list(idp_names, var_names))
spearman_matrix <- matrix(NA, nrow = length(idp_names), ncol = length(var_names),
                          dimnames = list(idp_names, var_names))

# Calculate correlation coefficients
for (idp in idp_names) {
  for (var in var_names) {
    pearson_matrix[idp, var] <- cor(df[[idp]], df[[var]], method = "pearson")
    spearman_matrix[idp, var] <- cor(df[[idp]], df[[var]], method = "spearman")
  }
}

# Melt the matrices for ggplot2
pearson_melt <- melt(pearson_matrix)
spearman_melt <- melt(spearman_matrix)

# Plot heatmaps
ggplot(data = pearson_melt, aes(x=Var2, y=Var1)) +
  geom_tile(aes(fill=value), color='white') +
  scale_fill_gradient(low = "blue", high = "red") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  labs(title = "Pearson Correlation Heatmap", x = "Control Variables", y = "IDP Variables", fill="Correlation")

ggplot(data = spearman_melt, aes(x=Var2, y=Var1)) +
  geom_tile(aes(fill=value), color='white') +
  scale_fill_gradient(low = "blue", high = "red") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  labs(title = "Spearman Correlation Heatmap", x = "Control Variables", y = "IDP Variables", fill="Correlation")

```
### visulization: control var ~ APOE_Status
```{r}
# Overlapping form of histograms
# Modify the functions to accept a title argument (overlapping histograms)
# plot_histograms <- function(var_name, df, title) {
#   p <- ggplot(df, aes_string(x = var_name, fill = "factor(ApoE_Status)")) +
#     geom_histogram(alpha = 0.5, position = "identity", bins = 30, aes(y = ..density..), color = "white") +
#     labs(x = var_name, fill = "APOE Status") +
#     scale_fill_manual(values = c("red", "blue")) +
#     theme_minimal() +
#     ggtitle(title)  # Add title to the plot
#   
#   print(p)
# }
# Modify the plot_histograms function to facet by ApoE_Status
plot_histograms <- function(var_name, df, title) {
  p <- ggplot(df, aes_string(x = var_name, fill = "factor(ApoE_Status)")) +
    geom_histogram(alpha = 0.5, position = "identity", bins = 30, aes(y = ..density..), color = "white") +
    labs(x = var_name, fill = "APOE Status") +
    scale_fill_manual(values = c("red", "blue")) +
    theme_minimal() +
    ggtitle(title) +  # Add title to the plot
    facet_grid(. ~ factor(ApoE_Status), scales = "free_y")  # Facet by ApoE_Status
  
  print(p)
}

plot_kdes <- function(var_name, df, title) {
  p <- ggplot(df, aes_string(x = var_name, color = "factor(ApoE_Status)")) +
    geom_density(size = 1.5) +
    labs(x = var_name, color = "ApoE Status") +
    theme_minimal() +
    ggtitle(title)  # Add title to the plot
  
  print(p)
}

df_AD <- df0 %>%
  filter(Amyloid_Status==1)
for (var_name in cont_control_names) {
  plot_histograms(var_name, df=df_AD, title="AD[+]")
  plot_kdes(var_name, df=df_AD, title="AD[+]")
}
```
```{r}
# Load necessary library
library(ggplot2)


# Loop through each variable in bi_control_names to create and display a bar plot
for (var in bi_control_names) {
  
  # Create the bar plot
  plot <- ggplot(df_AD, aes_string(x = "ApoE_Status", fill = var)) +
    geom_bar(position = "dodge", stat = "count") +
    labs(x = "ApoE Status", y = "Count", fill = var) +
    ggtitle(paste("Distribution of", var, "by ApoE Status")) +
    theme_minimal()
  
  # Display the plot
  print(plot)
  
}

```
```{r}
for (binary_var in bi_control_names) {
  tab <- table(df$ApoE_Status, df[[binary_var]])
  print(tab)
}
```


### visulization: control var ~ APOE dose
```{r}
control_names
```

```{r}
# Modify the functions to accept a title argument (overlapping histograms)
# plot_histograms <- function(var_name, df, title) {
#   p <- ggplot(df, aes_string(x = var_name, fill = "factor(ApoE_Status)")) +
#     geom_histogram(alpha = 0.5, position = "identity", bins = 30, aes(y = ..density..), color = "white") +
#     labs(x = var_name, fill = "APOE Status") +
#     scale_fill_manual(values = c("red", "blue")) +
#     theme_minimal() +
#     ggtitle(title)  # Add title to the plot
#   
#   print(p)
# }
# Modify the plot_histograms function to facet by ApoE_Status
plot_histograms <- function(var_name, df, title) {
  p <- ggplot(df, aes_string(x = var_name, fill = "factor(apoE4dose)")) +
    geom_histogram(alpha = 0.5, position = "identity", bins = 30, aes(y = ..density..), color = "white") +
    labs(x = var_name, fill = "apoE4dose") +
    scale_fill_manual(values = c("red", "blue", "green")) +
    theme_minimal() +
    ggtitle(title) +  # Add title to the plot
    facet_grid(. ~ factor(apoE4dose), scales = "free_y")  # Facet by ApoE_Status
  
  print(p)
}

plot_kdes <- function(var_name, df, title) {
  p <- ggplot(df, aes_string(x = var_name, color = "factor(apoE4dose)")) +
    geom_density(size = 1.5) +
    labs(x = var_name, color = "apoE4dose") +
    theme_minimal() +
    ggtitle(title)  # Add title to the plot
  
  print(p)
}

df_AD <- df0 %>%
  filter(Amyloid_Status==1)
for (var_name in cont_control_names) {
  plot_histograms(var_name, df=df_AD, title="AD[+]")
  plot_kdes(var_name, df=df_AD, title="AD[+]")
}
```
```{r}
df_AD %>%
  group_by(apoE4dose) %>%
  summarise(n=n())
```

```{r}
# Load necessary library
library(ggplot2)


# Loop through each variable in bi_control_names to create and display a bar plot
for (var in bi_control_names) {
  
  # Create the bar plot
  plot <- ggplot(df_AD, aes_string(x = "apoE4dose", fill = var)) +
    geom_bar(position = "dodge", stat = "count") +
    labs(x = "apoE4dose", y = "Count", fill = var) +
    ggtitle(paste("Distribution of", var, "by apoE4dose")) +
    theme_minimal()
  
  # Display the plot
  print(plot)
  
}

```
```{r}
for (binary_var in bi_control_names) {
  tab <- table(df$ApoE_Status, df[[binary_var]])
  print(tab)
}
```

## 2. AD[-]: compare nonAPOE4 vs APOE4
### 2.1 Visualization
```{r}
# Modify the functions to accept a title argument
plot_histograms <- function(var_name, df, title) {
  p <- ggplot(df, aes_string(x = var_name, fill = "factor(ApoE_Status)")) +
    geom_histogram(alpha = 0.5, position = "identity", bins = 30, aes(y = ..density..), color = "white") +
    labs(x = var_name, fill = "APOE Status") +
    scale_fill_manual(values = c("red", "blue")) +
    theme_minimal() +
    ggtitle(title)  # Add title to the plot
  
  print(p)
}

plot_kdes <- function(var_name, df, title) {
  p <- ggplot(df, aes_string(x = var_name, color = "factor(ApoE_Status)")) +
    geom_density(size = 1.5) +
    labs(x = var_name, color = "ApoE Status") +
    theme_minimal() +
    ggtitle(title)  # Add title to the plot
  
  print(p)
}

for (var_name in idp_names) {
  plot_histograms(var_name, df=df_noAD, title="AD[-]")
  plot_kdes(var_name, df=df_noAD, title="AD[-]")
}
```
```{r}
df_noAD %>%
  group_by(ApoE_Status) %>%
  summarise(n=n())
```

```{r}
formula <- as.formula(paste("ApoE_Status", "~", paste(idp_names, collapse = " + ")))
balance_table<-bal.tab(formula, data = df_noAD, disp = c("mean", "sd"), thresholds = 0.1)
print(balance_table)
```
Missing pattern:
```{r}
df1<-df_noAD[c("ID", 'Amyloid_Status',control_names,idp_names)]
missing_names<-colnames(df1)[apply(is.na(df1), 2, any)]

md.pattern(df1[,missing_names], rotate.names = TRUE)
colSums(is.na(df1[,missing_names])) #number of missing values
#rlv_names
```

### 2.2 regression: Idp ~ ApoE_Status
```{r}
# Initialize an empty list to store significant models
significant_models <- list()

# Loop through each variable in idp_names
for (idp in idp_names) {
  
  # Build the formula for lm
  formula <- as.formula(paste(idp, "~ ApoE_Status"))
  
  # Fit the linear model
  model <- lm(formula, data = df_noAD)
  
  # Check the significance of the Amyloid_Status coefficient
  coef_summary <- summary(model)$coefficients
  
  # The row corresponding to Amyloid_Status in coef_summary will be 2
  # Check if the p-value is less than 0.1
  if (coef_summary[2, 4] <= 0.1) {
    significant_models[[idp]] <- model
  }
}

# Now significant_models contains all models with a significant coefficient 
# for Amyloid_Status at a level of 0.1

```
```{r}
significant_models
summary(significant_models[[2]])#p=0.134
```
```{r}
model <- lm(dlt_ratio_sAPPb_sAPPa~ApoE_Status, data = df_AD)
summary(model)
```

### Multiple comparisons: bonferroni correction

```{r}
# Initialize an empty list to store p-values
p_values <- vector()

# Loop through each variable in idp_names
for (idp in idp_names) {
  
  # Build the formula for lm
  formula <- as.formula(paste(idp, "~ ApoE_Status"))
  
  # Fit the linear model
  model <- lm(formula, data = df_noAD)
  
  # Check the significance of the ApoE_Status coefficient
  coef_summary <- summary(model)$coefficients
  
  # The row corresponding to ApoE_Status in coef_summary will be 2
  # Store the p-value in p_values
  p_values <- c(p_values, coef_summary[2, 4])
}

# Now adjust the p-values using Holm-Bonferroni method
adjusted_p_values <- p.adjust(p_values, method = "fdr")#“holm”, “hochberg”, “hommel”, “bonferroni”, “BH”, “BY”, “fdr”, “none”

# Initialize an empty list to store significant models
significant_models <- list()

# Loop through the adjusted p-values and check if they are below 0.1
for (i in seq_along(adjusted_p_values)) {
  if (adjusted_p_values[i] <= 0.1) {
    
    # Build the formula for lm
    formula <- as.formula(paste(idp_names[i], "~ ApoE_Status"))
    
    # Fit the linear model
    model <- lm(formula, data = df_AD)
    
    # Store the model in significant_models
    significant_models[[idp_names[i]]] <- model
  }
}

# Now significant_models contains all models with a significant coefficient 
# for ApoE_Status at a level of 0.1, adjusted for multiple comparisons

```
```{r}

adjusted_p_values
p_values
```

```{r}
significant_models
#summary(significant_models[[1]])#p=0.134
```